---
title: "corona_analysis"
output:
  html_document:
    fig_width: 6
    fig_height: 4

knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file, 
  encoding=encoding,
  output_file=here::here(out_dir,"index.html")
 )
 })
    
---



```{r setup, include=FALSE}


# publish rmarkdown to github pages: 
#https://resources.github.com/whitepapers/github-and-rstudio/

# file.path(here::dr_here(), out_dir, 'index.html'))})
# file.path(here::dr_here(out_dir, 'index.html'), "docs", 'index.html')
# 
# here::dr_here()

# output: 
#   html_document:
#     keep_md: true
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>",
  rownames.print = 50,
  cols.print = 20,
   fig_width =  6,
  fig_height = 4)

```

```{r, echo=FALSE, warning=FALSE}
require(pacman)
p_load(tidyverse)
p_load(lubridate)
p_load(magrittr)
p_load(fpp3)
p_load(plotly)
p_load(slider)
p_load(glue)
```

## CONFIG
### data source: https://github.com/CSSEGISandData/COVID-19
```{r }
data_sources <-
  list(
 confirmed = 
   "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv",
 deaths = 
   "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"
  )

MIN_SAMPLE_SIZE <- 14
FORECASTING_DAYS <-  7
COUNTRY_STATE_OF_INTEREST <- 
  list("China: Hubei", "China: Non-Hubei", "Korea, South", 
       "Spain", "Italy", "Norway", 
       "Israel")
```



## load & tidy data
```{r, echo = FALSE, message=FALSE, warning=FALSE}

df_locations_of_with_restriction_date <- 
  list(
    "China: Hubei" = "2020-01-24", # full lockdown most population; source - https://en.wikipedia.org/wiki/2020_Hubei_lockdowns#Hubei
    "China: Non-Hubei"= "2020-02-10", # # outdoor restrictions  (during early Feb; dat selected by Beijing which was one of the last places); source - https://en.wikipedia.org/wiki/2020_Hubei_lockdowns#Hubei
    "Korea, South" =  "2020-02-20", # partial restrictions source: 23 March https://en.wikipedia.org/wiki/2020_coronavirus_pandemic_in_South_Korea
    "Spain" =  "2020-03-14", #source: https://en.wikipedia.org/wiki/2020_coronavirus_pandemic_in_Spain (22-3-2020 in medInt)
    "Italy" =  "2020-03-11",  #source: https://en.wikipedia.org/wiki/2020_coronavirus_pandemic_in_Italy#Nationwide_measures,
    "Norway" = "2020-03-12",
    "Israel" = "2020-03-16",  # https://en.wikipedia.org/wiki/2020_coronavirus_pandemic_in_Israel
    "France: France"= "2020-03-16", # source https://en.wikipedia.org/wiki/2020_coronavirus_pandemic_in_France
    "Switzerland" = "2020-03-20", # partial restrictions # https://en.wikipedia.org/wiki/2020_coronavirus_pandemic_in_Switzerland
    "Sweden" = "2020-03-22", # mild social distaning - source: medInt
    "Germany" =  "2020-03-22",  #source: https://en.wikipedia.org/wiki/COVID-19_related_curfews_and_lockdowns
    # "Iran" =  NA,
    "Netherlands: Netherlands" = "2020-03-23", # strict social distancing rule https://en.wikipedia.org/wiki/2020_coronavirus_pandemic_in_the_Netherlands
    "United Kingdom: United Kingdom" = "2020-03-23" # source: https://en.wikipedia.org/wiki/2020_coronavirus_pandemic_in_the_United_Kingdom
  ) %>% 
  enframe(name = "country_state", value = "restrict_date") %>% 
  mutate(restrict_date = restrict_date %>% map(as.Date)) %>% 
  unnest(restrict_date)


data <- 
  local({
    
    data_raw <- 
      localizebase::cache_eval(
        keys = Sys.Date(),
        expr =
          {
            data_sources %>%
            map(read_csv)
           })

    data_tidy <- 
      data_raw$confirmed %>%
      select(-c("Lat", "Long")) %>%
      pivot_longer(-c("Province/State", "Country/Region"), "date") %>%
      set_colnames(c("state","country", "date", "confirmed")) %>%
      filter(confirmed>=which.min(confirmed>=1)) %>% 
      mutate(date=as.Date(date, "%m/%d/%y")) 
  
    df_proc <- 
      data_tidy %>% 
      mutate(state = 
               ifelse((country == "China" & state != "Hubei"),"Non-Hubei", state)) %>% 
      mutate(country_state = 
               paste0(country, ifelse(is.na(state),"", paste0(": ",state)))) %>%
      group_by(country, country_state, date) %>% # sum non Hubei provinces together
      summarise(confirmed = sum(confirmed)) %>% 
      ungroup() %>% 
      left_join(df_locations_of_with_restriction_date, 
                by = "country_state") %>% 
      group_by(country_state) %>%
      arrange(date) %>%
      mutate(
        new_cases = pmax(confirmed - lag(confirmed),0),
        r = confirmed/lag(confirmed),
        r = ifelse(is.infinite(r) | is.nan(r) | is.na(r), 1, r)
      ) %>% 
      mutate(mean_new_cases = slider::slide_dbl(new_cases, mean, .before=3),
             mean_r = slider::slide_dbl(r, mean, .before=1, .after=1)) %>% 
      mutate(
        # date_case100 = min(date[confirmed >= 100]),
        # date_peak = min(date[which.max(mean_r)]),
        date0 = min(date[date >= restrict_date
          # date >= date_peak + 7
          #date >= date_case100
          ]),
        dateEnd = min(date[date > date0 & mean_new_cases == 0 ]),
        # n_row = row_number(),
        # min_nday = min(n_row[date == date0]),
        # max_nday = max(n_row) - 1,
        # is_date0na = is.na(as.POSIXlt(date0)),
        # n_day = n_row - min(n_row[date == date0])
        # n_day = ifelse(is_date0na, # now day0 found - non of the dates meets the cretiria
        #                        n_row - max(n_row) - 1,
        #                        n_row - min(n_row[date == date0]))

       ) %>% 
      ungroup() %>% 
      arrange(country_state, date)
    
    df_clean <- 
      df_proc %>% 
      filter(country_state %in% COUNTRY_STATE_OF_INTEREST) %>% 
      select(country, country_state,date, restrict_date, everything())
    
    res <- 
      list(
        df_proc = df_proc,
        df_clean = df_clean
      )
  })


glue::glue("date range: 
           [{min(data$df_proc$date)} - {max(data$df_proc$date)}]")
```


## daily % change in total confirmed cases:  
## (1) trend extrapolation for next 7 days for selected countries (precede outbreak in IL), and
## (2) modeling various scenarios for Israel for the next 7 days (e.g. if its confirmed cases dynamic will be similar to Iran or S.Korea).
```{r }
# fit exponentinal smoothing model for region
model_data <- 
  local({
    
    tsbl <-
      data$df_clean %>% 
      mutate(r = ifelse(date == date0, 1, r)) %>% 
      # filter(n_day>=0) %>% 
      as_tsibble(key = c("country","country_state"),
                 index = date)
    
    ETS_model_fitted <- 
      tsbl %>% 
      model(ETS(r ~ error("A") + trend("A") + season("N"),  opt_crit = "mae")) 
    # !!rlang::sym(TARGET_VAR)
    
    interpolation <- 
      ETS_model_fitted %>% 
      augment() %>% 
      as_tibble() %>% 
      select(
        country,
        country_state, 
        date,
        r_model = .fitted)
    
     extrapolation <- 
       ETS_model_fitted %>% 
       forecast(h = FORECASTING_DAYS) %>%
       as_tibble() %>% 
       select(country,
              country_state, 
              date,
              r_model = r)
     
     fitted_data <- 
       bind_rows(interpolation = interpolation,
                 extrapolation = extrapolation,
                 .id = "fit_type") %>% 
       mutate(r_model = pmax(r_model,1))
     
     model_data <- 
       fitted_data %>% 
       full_join(data$df_clean, 
                 by = c("country","country_state", "date")) %>% 
       mutate(prc_change_in_total_cases_model = 100*(r_model - 1)) %>%
      group_by(country_state) %>% 
       arrange(date) %>% 
       # fill missing date for extrapulated value 
       mutate(
         # date_latest = date[which.max(date)], #only interpulated data has dates after the join
         # n_day_latest = max(n_day[which.max(date)]),
         confirmed_latest = max(confirmed[which.max(date[fit_type ==  "interpolation"])]) #max(confirmed[which.max(date)]),  
         # confirmed_first = min(confirmed[which.min(date)]),
         # date = date_latest + days(n_day-n_day_latest)
         ) %>% # add date to extrapolated values
       ungroup() %>%
       # calc confirmed case by model r for the extrapolated entries
       (function(data){
         extrapolation <-
           data %>%
           filter(fit_type == "extrapolation") %>%
           group_by(country_state) %>%
           arrange(date) %>%
           mutate(
             #confirmed = confirmed_latest,
                  confirmed_model = # calc confirmed cases by r_model
                    Reduce(init = max(confirmed_latest),
                           x = lead(r_model), #lead(r_model),
                           f = prod, 
                           accumulate = T)[-1]
                  # confirmed = NA_integer_
                  ) %>% 
           ungroup() 
         data %>%
           filter(fit_type == "interpolation") %>%
           bind_rows(extrapolation)
        }) %>% 
       select(
         country,
         country_state, 
         # n_day, 
         date,
         date0, restrict_date, 
         new_cases, confirmed, confirmed_model,
         r, fit_type, r_model, prc_change_in_total_cases_model
         # date_peak
       )
  })

```


# log(Total confirmed) cases by day since case 1.
```{r, fig.width=4, fig.height=2}
local({
  
  data_for_plot <- 
    model_data %>% 
    group_by(country_state) %>% 
    mutate(n_day = difftime(date, min(date), units = "days") %>% as.integer()) %>% 
    mutate(n_day_restrict = min(n_day[date == restrict_date], na.rm = T),
           country_state_restrict_date = paste0(country_state,"; social restriction: ", restrict_date)) %>% 
    ungroup() %>% 
    filter(fit_type == "interpolation") 
  
  plot <- 
    ggplot(data = data_for_plot,
           aes(x = n_day,
               y = log(confirmed), 
               date = date, 
               color = country_state,
               r_model = round(r_model,3)
               ),
           show.legend = FALSE) +
    geom_point(alpha = 0.6, color = "black", show.legend = FALSE)+
    geom_line(alpha = 0.6,show.legend = FALSE)+
    ggtitle("Total confirmed COVID-19 cases by country") +
    geom_vline(aes(xintercept = n_day_restrict, date= date),
               show.legend = FALSE)  +
    geom_text(mapping = aes(label = "social restriction",
                            x = n_day_restrict,
                            y = -Inf
                            ),
              angle = 90, hjust = -0.5, nudge_x = 1,
              size = 1.5,
              show.legend = FALSE)+
    theme(legend.position='none')+
     facet_wrap(~country_state_restrict_date, 
               scales = "free") 
  
  suppressWarnings(print(plot))
  
  ggp_build <- plotly_build(plot)
  ggp_build$layout$height = 800
  ggp_build$layout$width = 1200
  
  print(ggp_build)
})
```



# Log(Total confirmed cases) by Date
```{r, fig.width=4, fig.height=2.1}
local({
  
  data_for_plot <- 
    model_data %>% 
    group_by(country_state) %>% 
    mutate(n_day_since_rest = difftime(date, restrict_date, units = "days") %>% as.integer() ,
    # mutate(n_day_restrict = min(n_day[date == restrict_date], na.rm = T),
           country_state_restrict_date = paste0(country_state,"; social restriction: ", restrict_date)
           ) %>% 
    ungroup() %>% 
    filter(fit_type == "interpolation") 
  
  plot <- 
    ggplot(data = data_for_plot,
           aes(x = n_day_since_rest, 
               y = confirmed, 
               n_day_since_rest = n_day_since_rest,
               # n_day = n_day, 
               # restrict_date = restrict_date,
               # country_state_restrict_date = country_state_restrict_date,
               color = country_state,
               r_model = r_model)
           ) +
    geom_point(alpha = 0.6, color = "black") +
    geom_line(alpha = 0.8) +
    ggtitle("Log(confirmed) by num days since social restriction") +
    # geom_vline(aes(xintercept = 0),
    #            show.legend = FALSE)  +
    # geom_text(aes(label = "restrict_date",
    #                         x = 0,
    #                         y = -Inf),
    #           angle = 60, hjust = -0.1,nudge_x = 1,
    #           show.legend = FALSE) +
    # ylab("total confirmed cases") +
    xlab("days since social restriction") +
    theme(legend.position='none')+
    scale_y_log10()+
    facet_wrap(~country_state_restrict_date,
               scales = "free")
  
  suppressWarnings(print(plot))

  ggplotly(plot)
  ggp_build <- plotly_build(plot)
  ggp_build$layout$height = 800
  ggp_build$layout$width = 1400

  print(ggp_build)
})
```



## let's examine rate - % change in *total* cases by day 
### (if % change = 20% in a given day ==>  total patients grew by 20% relative to the previous day).  
### the dashed line show extrapulation for 7 days forward.
```{r, fig.width=10, fig.height=8}
# daily percent change in *total* confirmed cases
local({
  
  data_for_plot <- 
    model_data %>% 
    group_by(country_state) %>% 
    mutate(n_day_since_rest = difftime(date, restrict_date, units = "days") %>% as.integer() ,
           country_state_restrict_date = paste0(country_state,"; social restriction: ", restrict_date),
           min_date = min(date),
           max_date = max(date)
    ) %>% 
    ungroup() %>% 
    filter(fit_type == "interpolation") 

  plot <- 
    ggplot(data = data_for_plot,
           aes(x = n_day_since_rest,
               y = 100*(r-1), 
               color = country_state,
           ),
           show.legend = FALSE) +
    geom_point(alpha = 0.4, 
               color = "black") +
    geom_line(
      data = data_for_plot %>% filter(fit_type=="interpolation"),
      aes(y=prc_change_in_total_cases_model, color = country_state),
      show.legend = FALSE,
      alpha = 0.8, size = 1.0) +
    geom_line(
      data = data_for_plot %>% filter(fit_type == "extrapolation"),
      aes(y = prc_change_in_total_cases_model, 
          color = country_state),
      show.legend = FALSE,
      linetype = "dashed"
    ) +
    ggtitle("% change in total cases") + #glue("date range [{min_date} - {max_date} ]")
    ylab("% change in total cases") +
    facet_wrap(~country_state_restrict_date,
               scales = "free")
  
  suppressWarnings(print(plot))
  
})

```


# forecast scenarios total number of cases for Israel based on Israel itself and other countries such as S.Korea, Italy, Iran
```{r, fig.width=8, fig.height=3}
df_latest_day_IL <- 
  model_data %>% 
  filter(country_state == "Israel", 
         fit_type == "interpolation") %>% 
  top_n(wt = date, n = 1)

print(glue("Israel at {df_latest_day_IL$date}: 
            r_model {round(df_latest_day_IL$r_model,3)} 
            confirmed cases: {df_latest_day_IL$confirmed} "))

# generate forecast for IL based on each countrty (calc 7 days fwd r_model)
ts_forecast <- 
  model_data %>% 
  filter(country_state != "Israel") %>% 
  filter(date >= date0) %>% 
  group_by(country_state) %>% 
  arrange(date) %>% 
  mutate(
    abs_diff = abs(r_model - df_latest_day_IL$r_model),
    min_diff_date = date[which.min(abs_diff)],
    dateStart = min_diff_date,  #note that this is "today" not a forecasted day
    dateEnd = dateStart + FORECASTING_DAYS) %>% 
  filter(date >= dateStart, date <= dateEnd) %>%
  filter(min(abs_diff)<=0.1) %>% 
  filter(n() == FORECASTING_DAYS + 1) %>% 
   mutate( # fill mising dates
    date_il = df_latest_day_IL$date + lubridate::days(0:FORECASTING_DAYS),
    confirmed_forecast_isr = # predict IL by other cointries
      Reduce(init = df_latest_day_IL$confirmed,
             x = lead(r_model), 
             f = prod, accumulate = T) %>% head(-1)
    ) %>% 
  # mutate(r_model_il = confirmed_forecast_isr/lag(confirmed_forecast_isr),
  #        prc_change_in_total_cases_model_il = 100*(r_model_il-1)) %>%  # recalc according to the confirmed forecast_isr
  ungroup() %>% 
  arrange(country_state, date) %>% 
  select(country_state, date, dateStart, dateEnd,
         r_model, prc_change_in_total_cases_model, date_il, confirmed_forecast_isr)

ts_forecast_with_il <- 
  ts_forecast %>% 
  # select(country_state, date, dateStart, dateEnd, date_il, prc_change_in_total_cases_model, confirmed_forecast_isr) %>% 
  bind_rows(model_data %>% 
              filter(country_state == "Israel") %>% 
              select(country_state, 
                     date,
                     date_il = date, 
                     prc_change_in_total_cases_model,
                     confirmed_forecast_isr = confirmed_model)
  ) %>% 
  select(country_state, date,dateStart, dateEnd, prc_change_in_total_cases_model, 
         date_il, confirmed_forecast_isr)
         #,r_model_il, prc_change_in_total_cases_model_il)



# layers: (1) regions raw data points (2) interpulation of raw points + (3) extrapulation (dashed line) + (4) fat lines for date-interval (7 days) used to predict IL
local({
  plot <- 
    ggplot(model_data %>% semi_join(ts_forecast_with_il, "country_state") %>% 
             filter(country_state != "Israel",fit_type == "interpolation"),
           aes(x = date,y = prc_change_in_total_cases_model, color = country_state),
           show.legend = FALSE) +
    geom_point(alpha = 0.1, color = "black") + # raw data
    geom_line( # smooting of raw data
      show.legend = FALSE,
      alpha = 0.5, size = 0.9) +
    geom_line( #forecast of raw
      data = model_data %>% semi_join(ts_forecast_with_il, "country_state") %>% 
        filter(country_state != "Israel", fit_type == "extrapolation"),
      show.legend = FALSE,
      linetype = "dashed"
      ) +
    geom_vline(data = ts_forecast_with_il %>% filter(country_state != "Israel"), 
               aes(xintercept = dateStart))+
    geom_vline(data = ts_forecast_with_il %>% filter(country_state != "Israel"), 
               aes(xintercept = dateEnd))+
    ggtitle("date range selected for Israel forecasting") +
    ylim(0, 70) +
    ylab("% change in total cases") +
    # scale_x_date(date_breaks = "3 weeks", date_labels =  "%d-%b") +
    facet_wrap(~country_state, scales = "free") 
  
  suppressWarnings(print(plot))
})
```


# forecast IL total cases over time by ref. regions
```{r}

local({

  plot <-  
    ggplot(
      data = 
        model_data %>% 
        filter(country_state == "Israel", fit_type == "interpolation"),
      mapping = aes(x = date, y = confirmed, color = country_state),
      show.legend = FALSE) +
    geom_point(alpha = 0.3, 
               color = "black") + # raw data
    geom_smooth(se=F, method = "gam") +
    geom_line(# smooting of raw data
      data = model_data %>% 
                 filter(country_state == "Israel", fit_type == "extrapolation"),
      mapping = aes(x = date, y = confirmed_model),
       alpha = 0.5, size = 0.9,
      linetype = "dashed"
      )  +
    geom_line(
      data = ts_forecast_with_il, #%>% filter(country_state != "Israel"),
      aes(x = date_il, y = confirmed_forecast_isr, color = country_state),
      linetype = "dashed",
      size = 1
    ) +
    ggtitle("Israel - total confirmed cases forecasting 7 days fwd by ref. countries") +
    ylab("total confirmed cases") 
  
  suppressWarnings(print(plot))
})
```



# IL total cases over time (with 7 days extrapulation) + prediction base on ref. countries.
```{r}

local({
  
  plot <- 
    ggplot(model_data %>% 
             filter(country_state == "Israel", fit_type == "interpolation"),
           aes(x = date,y = prc_change_in_total_cases_model),
           show.legend = FALSE) +
    geom_line( # smooting of raw data
      show.legend = FALSE,
      alpha = 0.5, size = 0.9) +
    geom_point(data = model_data %>% filter(country_state == "Israel", fit_type == "interpolation"),
              alpha = 0.3, color = "black") + # raw data
    geom_line( # data used for IL prediction
      data = ts_forecast_with_il, #%>% filter(country_state != "Israel"),
      aes(x = date_il, y = prc_change_in_total_cases_model, color = country_state),
      linetype = "dashed",
      size = 1
    ) +
    ggtitle("Israel - forecasting of daily change in *total confirmed cases* by ref. countries") +
    ylim(0, 100) +
    ylab("% change in total cases") 
  
  suppressWarnings(print(plot))
})
```


